
\documentclass{ClassSettings}

\begin{document}

% UoB guidlines:
%
% Preliminary pages
% 
% The five preliminary pages must be the Title Page, Abstract, Dedication
% and Acknowledgements, Author's Declaration and Table of Contents.
% These should be single-sided.
% 
% Table of contents, list of tables and illustrative material
% 
% The table of contents must list, with page numbers, all chapters,
 % sections and subsections, the list of references, bibliography, list of
% abbreviations and appendices. The list of tables and illustrations
% should follow the table of contents, listing with page numbers the
% tables, photographs, diagrams, etc., in the order in which they appear
% in the text.
% 
\frontmatter
\pagenumbering{roman}


LZ SR1 WIMP result is expected this month (June 2022), with an EFT result expected imminently after; in the same energy and extended.
This document is the state of thesis as of 10-June-2022.
Status of analysis;
\begin{itemize}
    \item \textbf{Simulations}: GPU, cavern-gamma and S2 map work completed. 85-90\% written
    \item \textbf{OD construction and simulation}: analysis complete, 60\% written.
    \item \textbf{OD analysis}: Core-analysis complete; OD bg rate and Veto eff and particle discrimination. Background fitting incomplete. 35\% written.
    \item \textbf{EFT}: 1000-day projected limits done. SR1 limits possible within thesis time-line. Expect limits 1-month after WIMP search result. 2\% written.
\end{itemize}

\clearemptydoublepage
%
\input{Title_Pages/title}
\clearemptydoublepage
%
\input{Title_Pages/abstract}
\clearemptydoublepage
%
\input{Title_Pages/dedication}
\clearemptydoublepage
%
\input{Title_Pages/declaration}
\clearemptydoublepage
%
\renewcommand{\contentsname}{Table of Contents}
\maxtocdepth{subsection}
\tableofcontents*
\addtocontents{toc}{\par\nobreak \mbox{}\hfill{\bf Page}\par\nobreak}
\clearemptydoublepage
%
\listoftables
\addtocontents{lot}{\par\nobreak\textbf{{\scshape Table} \hfill Page}\par\nobreak}
\clearemptydoublepage
%
\listoffigures
\addtocontents{lof}{\par\nobreak\textbf{{\scshape Figure} \hfill Page}\par\nobreak}
\clearemptydoublepage
%
%
% The bulk of the document is delegated to these chapter files in
% subdirectories.
\mainmatter
%
\input{Chapters/main}
\clearemptydoublepage

\backmatter
%\bibliographystyle{siam}
%\refstepcounter{chapter}
\bibliographystyle{IEEEtran}
\bibliography{References/References.bib,References/Opticks.bib,References/LZ_References.bib,References/GdLS_Physics.bib,References/EFT_References.bib,References/DarkMatter_Evidence.bib,References/OD_Particle_ID.bib,References/Calibration_References.bib,References/Construction.bib,References/DM_Theory_References.bib}
\clearemptydoublepage
%
% Add index
%\printindex
%   
\end{document}
