\chapter{Analysis of the Outer Detector} \label{chap:analysis_of_the_od}

\par
In this chapter, the calibration of the OD is described.
This forms the foundations for assessing if the requirements set out in \autoref{tab:veto_requirements} are achieved.
These requirements are: that the neutron veto efficiency is at least 95\% and that no more than 5\% of the WIMP search time is vetoed.
For ascertaining an appropriate veto window, a study on the OD backgrounds is performed. 
This is followed by the results from a neutron calibration.
%\par
%Finally a particle discrimination method is suggested, which in future may allow for the OD to be used as a more intelligent veto.


\input{Chapters/Analysis_OuterDetector/od_comissioning}
\clearpage
\input{Chapters/Analysis_OuterDetector/od_backgrounds}
\clearpage
\input{Chapters/Analysis_OuterDetector/neutron_inefficiency}

%\input{Chapters/Analysis_OuterDetector/particle_discrimination}

\subsubsection*{Summary}
\par
In this chapter, the backgrounds in the OD were measured and fitted.
Most backgrounds were within error of what was expected but there was an elevated rate of ${}^{210}$Po which has been linked to cavern air entering the acrylic tanks during construction.
The neutron veto efficiency was measured as 84.6\%, which is lower than the 91.0\% expected from simulations and the 95\% required by the design requirements.
The most likely cause of this was determined to be the neutrons taking a longer time to reach the GdLS.
The dominant factor appears to be water ingress into the foam which surrounds the OCV.
Neutrons thermalise in this water, causing them to take longer to reach the GdLS and can also be captured there.