%
% file: dedication.tex
% author: V?ctor Bre?a-Medina
% description: Contains the text for thesis dedication
%

\chapter*{Acknowledgements}
\begin{SingleSpace}
\par
It would not have been possible to get to the stage where this thesis is written without the help of numerous people.
I will not be able to mention everyone in such a brief acknowledgement but if I worked with you over the last 4-years I'd like you to know it was a pleasure.
\par
Firstly I'd like to thank my parents for instilling a love of education and physics in me. I appreciate the fact that you continued to have faith in me even it was difficult to (my 11-plus, GCSE and A-level results being just a few examples).
Without your nurture I wouldn't be who I am today.
And to my brother Ben.
You provided me with the distraction from physics whenever I needed it, particularly when I couldn't bare to write anymore.
\par
To my supervisor, Henning, thanks for taking a chance on me four years ago and giving me the opportunity to work on a collaboration full of bright minds. 
Your support and guidance has been greatly valued.
To my friends who I started with in Bristol, Lakshan, Nikos and David, we learned how to navigate the world of a PhD student together and had a few drinks along the way. 
Thanks for making my first year so enjoyable.
We started together and now we're finishing together!
\par
Ben and Luke, both of you helped me a tremendous amount in my first year when I had no idea what I was going (and not convinced I have any idea now either).
I was still be bashing my head trying to work out how to use Oracle Cloud Computing without you guys.
\par
Although I started my PhD in Bristol, I was fortunate enough to travel to and work in several destinations in the United States, meeting some incredible people on the way. Working with the Michigan group has been a great pleasure. 
I would like to thank Bj\"{o}rn for being so welcoming to me in both Boston and Michigan.
You gave me the resources and opportunity to really embed myself within LZ and set me up for a career in academia.
I'd also like to acknowledge and thank my fellow post-graduates, Ewan, Luke, Mike, Harvey, Maris and Chami. 
I'll remember our conversation on physics, travel and everything in between for a long time. 
I am excited to see your careers progress and hope to work together again in the near future.
\par
For my time on-site I cannot thank Derek and Ryan enough. As you know at times it was challenging with working night shifts, barely seeing any sunlight and being separated from the world because of the pandemic. There are no two other people I would rather have spent this time with. I learned many valuable skills during this time, many of them from the two of you. 
\par
To my friends James, Rich, Tim and JJ.
Thank you for keeping me sane in the first few months of the pandemic.
Having our daily gaming sessions was a wonderful distraction.
\par
It would not have been possible to have completed my PhD without the support of Keith.
I was always excited to share the latest of the project with you.
\par
Finally, to Anika.
Your love and support during my PhD make it possible.
I wouldn't have gotten very far without you (I'd definitely have been homeless when I moved the the States).
You are my rock, and you kept me going when I thought I couldn't go on.
\end{SingleSpace}
\clearpage