\chapter{Dark Matter - The Unknown Mass}

\par
For thousands of years the humans have tried to understand the Universe that we inhabit.
The majestic beauty of the sky has been one of most intriguing things - with our knowledge developing from believing we are the centre to realising that we are a tiny part of the Universe and the Universe is a dynamic and ever changing landscape (all-be-it very slowly on our time-scale).

\par
However, as our understanding increasing, we have realised that we may in-fact understand significantly less of the Universe and it's components that we thought.

What's more amazing is that technological improvements have allowed for a much greater understanding of the Universe we live in.
This has included the first image of a black hole, and the observations of gravitation waves from merging neutron stars. 
TODO add cites.
Though our understanding of what is truly the makeup of our Universe is tiny - at 5\%. 
This leaves a massive amount of unknown.

For nearly a century, cosmological observations have given results that do not fit within the standard model of particle physics (SM), which though can predict most experimental results, cannot account for gravity and conflicts with general relativity. 
This leads for the need to go beyond the SM. It is known that the SM can only account for 5\% of the energy and matter in the Universe. 
The other parts, known as dark matter (DM) and dark energy due to their non-luminous properties are taken into account in the $\Lambda$-CDM model of the Universe which is the most inline with cosmological observations.
It is also interesting to note that extensions to the SM to solve problems such as the hierarchy problem naturally lead to DM.


\par
As technology has advanced we have gained the ability to peer deeper and deeper into space.

\par
In this chapter we summarise the evidence for this unexplained part of our universe.
This is followed by a selection of candidates that have been hypothesised to be dark matter, and the way in which they are being searched for.

\input{Chapters/DarkMatter/evidence}

\input{Chapters/DarkMatter/candidates}
