\chapter{The LUX-ZEPLIN Experiment}
\label{sec:lz_detector_chapter}
\par
As discussed in the previous chapter, there is a reasonable expectation that if dark matter is comprised at least in part of WIMPs, then WIMP-nucleon scattering should be detectable.
The question then becomes, what is the optimal detector to detect this scattering?
As the WIMP-nucleon scattering is a rare event, a low background detector design is required.
Regardless of what the detector is made from, there will always be some backgrounds, most commonly $\beta$s and $\gamma$s that need to be categorised and removed.
\par
In this chapter, the concepts of xenon-based dark matter detectors are introduced.
This is then followed by the presentation of the LUX-ZEPLIN dark matter experiment upon which the analyses in this thesis are based.
This is not the only type of direct dark matter detector, and for a description of other detector designs, the reader is directed to \cite{direct_detector_designs_ref}.

\input{Chapters/LUXZEPLIN/wimps_in_tpcs}

\clearpage
\input{Chapters/LUXZEPLIN/lz_detector}

%\input{Chapters/LUXZEPLIN/lz_event}

\subsubsection*{Summary}
\par
This chapter began with an overview of Xenon physics and how energy deposits can be observed.
This was used to explain how Time-Projection Chambers (TPC) can be used to search from WIMP-nucleon recoils.
The direct dark matter experiment LUX-ZEPLIN, which has at its core a TPC, was then summarised.
Each of the sub-detectors was described along with the backgrounds which are observed by the experiment.
An overview of the dark matter search strategy was then given.
Finally, an overview of the simulation and analysis chain of the LUX-ZEPLIN experiment was presented.